---
title: "Generate Package Website with pkgdown and Github Pages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generate Package Website with pkgdown and Github Pages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PkgTestR)
```

## Set up Git and Github 

Develope a package, that includes functions with ROxygen2 documentation and also vignette files. Then, use functions from *usethis** to automate various git and github related tasks.

First, initialize the project folder as a git folder, and then commit all existing non-gitignored files. See the *.gitignore* file for which documents are ignored. 

```{r, eval = FALSE}
usethis::use_git(message = "Initial commit")
```

nce per package:
• use_git() — Activate git
• use_github() — Connect to GitHub
• use_github_actions() — Set up automated checks

## Testing

```{r, eval = FALSE}
usethis::use_pkgdown_github_pages()
```
